\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2021-12-20 main.Rmd
more writing and updates on the interactive graphic section

With a larger dataset, interactive graphics can be helpful to explore these data from multiple perspective. Two examples are presented here to show how interactivity can be used to explore the climate data across Australia. Again, `weatherdata::climate_full` will be used and we hope to include more than one time variable here. The difference between the maximum and minimum temperature of the day, the diurnal range, can be directly calculated from the data and it can have large impact on the wine industry (lol this is going far). This would be useful to compare the temperature difference across stations and to compare how this temperature difference differs throughout the year.

Temperature difference is a time-wise variable and should be computed in the long form cubble. The variables are averaged over the 2016 to 2020 period by month to stabilise the result.

```{r eval = FALSE, echo = TRUE}
res <- weatherdata::climate_full %>%
  # filter(!id %in% c("ASN00067033", "ASN00072091", 
  #                   "ASN00059040", "ASN00097053")) %>%
  stretch() %>%
  mutate(month = lubridate::month(date),
         diff = tmax - tmin) %>%
  group_by(month) %>%
  summarise(tmax = mean(tmax, na.rm = TRUE),
            diff = mean(diff, na.rm = TRUE)) %>%
  ungroup(month)

@@

Several variables can be useful when comparing the seasonal effect of the temperature differences, for example, the difference of average temperature difference between summer and winter can tell [...]. This is a station-wise measure and hence computed in the nested form.

```{r eval = FALSE, echo = TRUE}
res <- res %>% 
  tamp() %>%
  mutate(
    diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
      pull(diff) %>% mean(na.rm = TRUE),
    diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
      pull(diff) %>% mean(na.rm = TRUE),
    diff_year = diff_winter - diff_summer
  )
```

The result from the above manipulation can be directly supplied to make interactive plots and Figure \ref{fig:interactive-linking} is an example made with \pkg{plotly} and \pkg{crosstalk}. For example, the second row results from selecting the highest mean maximum temperature in July and this corresponds to Kalumburu where there is a larger winter (June to August) temperature difference than summer (December to February). The third row selects one of the stations (Birdsville Airport) with a large average temperature difference in the interior desert region on the map and the linked view from the time series shows that this maximum and minimum temperature difference is constant throughout the year. The last row selects the lowest maximum temperature in July on the temperature band and surprisingly, the corresponding station doesn't fall into the Tasmania island but Mountain Hotham in Victoria.

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia. Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively. The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia. The html version can be found at ...", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-mid.png"))
knitr::include_graphics(here::here("figures/linking-south.png"))
```

Cubble is also compatible with \pkg{leaflet} and with leaflet extension package \pkg{leafpop}, the temperature band can be displayed upon clicking on the station on the map. Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia. With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature throughout the year. With a slightly higher latitude than the mountain Hotham, flat land in Melbourne airport shows a less extreme temperature band. Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows a uniformly low and thin temperature band.

```{r interactive-popup, echo = FALSE, out.width="50%", out.height="30%", fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia. The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window. The html version can be found at ...", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-north.png"))
knitr::include_graphics(here::here("figures/popup-vic.png"))
knitr::include_graphics(here::here("figures/popup-tas.png"))
```

>>>>

To visualise climate variables from all the stations in Australia, interactive graphics can be helpful. Here a plotly and a leaflet example is given to explore the yearly diurnal temperature range pattern in Australia. This would be useful to see how the average temperature range differs in various latitudes in Australia, whether there's a summer or winter effect, any coastline, mountain or desert topological considerations that would affect the temperature.

Again, `weatherdata::climate_full` will be first processed to produce summarised monthly maximum and minimum temperature across 2016 to 2020. Then a shared data with the same group is created from the nested and long cubble to enable the cross-talking via the common column `id`. The map and temperature band is created separately with the two shared data objects and combined together with `crosstalk::bscols`. A sketch of the code to produce Figure \ref{fig:interactive-linking} is presented below:

```{r eval = FALSE, echo = TRUE}
# data processing
clean <- weatherdata::climate_full %>% ...

# created grouped instance for crosstalk
nested <- clean %>% SharedData$new(~id, group = "cubble")
long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")

# create the map with SharedData nested, the time series with SharedData long
p1 <- nested %>% ...
p2 <- long %>% ...

# Combine p1 and p2
crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)

@@

In Figure \ref{fig:interactive-linking}, the first row shows the default plot and when a selection is made, relevant component from both plots will be highlighted. For example, in the second row, the highest mean maximum temperature in July is selected on the time series and this corresponds to Kalumburu on the map where there is a larger diurnal temperature range in winter (June to August) than summer (December to February). The third row selects a station (Birdsville Airport) in the interior desert region that has a large average temperature difference on the map and the linked view shows a wide but constant diurnal temperature range throughout the year. The last row selects the lowest maximum temperature in July on the right plot and surprisingly, this links to the Thredbo airport at the boarder of Victoria and New South Wales, rather than any station in the Tasmania island!

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the mean daily temperature difference averaged by month from 2016 to 2020 at various locations in Australia. Crossed linking between the map and the temperature band is implemented in both directions for exploring the temperature pattern interactively. The first row is the general appearance of the graph and the next three rows are examples from various selection to show different temperature patterns in Australia. The html version can be found at ...", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-mid.png"))
knitr::include_graphics(here::here("figures/linking-south.png"))
```

With leaflet popup, the temperature range can be displayed as a small plot upon clicking on the station on the map. This would require first creating all the popup plot separately in a vector and then add these plots to a leaflet map with `leafpop::addPopupGraphs()`:

```{r eval = FALSE, echo = TRUE}
# create subplots for each station
df_id <- unique(clean$id)
p <- map(1:length(df_id), function(i){
  dt <- clean %>% filter(id == df_id[i])
  dt %>% ...
})

# inset subplots as a popup in leaflet
leaflet(clean) %>% 
  addTiles() %>% 
  addCircleMarkers(group = "a", ...) %>% 
  leafpop::addPopupGraphs(graph = p, ...)

Figure \ref{fig:interactive-popup} shows more examples of temperature pattern that can be found in Australia. With the same latitude as Kalumburu but in the Pacific coastline, Cooktown airport shows a similarly high but constant temperature range throughout the year. With a slightly higher latitude than the Thredbo airport , Melbourne airport shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in the north-west tip of the Tasmania island shows low and narrow temperature band, constantly betwee 10 and 20 degree throughout the year.

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="30%", fig.retina = 2, dpi = 300, fig.cap = "Screenshots of the diurnal temperature range averaged by month from 2016 to 2020 at various locations in Australia. The leaflet popup allows users to click on a single station for its temperature band as a small plot in the popup window. The html version can be found at ...", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-north.png"))
knitr::include_graphics(here::here("figures/popup-vic.png"))
knitr::include_graphics(here::here("figures/popup-tas.png"))
```

### 2021-12-20 main.tex
more writing and updates on the interactive graphic section

With a larger dataset, interactive graphics can be helpful to explore
these data from multiple perspective. Two examples are presented here to
show how interactivity can be used to explore the climate data across
Australia. Again, \texttt{weatherdata::climate\_full} will be used and
we hope to include more than one time variable here. The difference
between the maximum and minimum temperature of the day, the diurnal
range, can be directly calculated from the data and it can have large
impact on the wine industry (lol this is going far). This would be
useful to compare the temperature difference across stations and to
compare how this temperature difference differs throughout the year.

Temperature difference is a time-wise variable and should be computed in
the long form cubble. The variables are averaged over the 2016 to 2020
period by month to stabilise the result.

\begin{CodeChunk}
\begin{CodeInput}
R> res <- weatherdata::climate_full %>%
+   # filter(!id %in% c("ASN00067033", "ASN00072091", 
+   #                   "ASN00059040", "ASN00097053")) %>%
+   stretch() %>%
+   mutate(month = lubridate::month(date),
+          diff = tmax - tmin) %>%
+   group_by(month) %>%
+   summarise(tmax = mean(tmax, na.rm = TRUE),
+             diff = mean(diff, na.rm = TRUE)) %>%
+   ungroup(month)
\end{CodeInput}
\end{CodeChunk}

Several variables can be useful when comparing the seasonal effect of
the temperature differences, for example, the difference of average
temperature difference between summer and winter can tell {[}\ldots{]}.
This is a station-wise measure and hence computed in the nested form.

\begin{CodeChunk}
\begin{CodeInput}
R> res <- res %>% 
+   tamp() %>%
+   mutate(
+     diff_winter = ts %>% filter(month %in% c(6, 7, 8)) %>%
+       pull(diff) %>% mean(na.rm = TRUE),
+     diff_summer = ts %>% filter(month %in% c(12, 1, 2)) %>%
+       pull(diff) %>% mean(na.rm = TRUE),
+     diff_year = diff_winter - diff_summer
+   )
\end{CodeInput}
\end{CodeChunk}

The result from the above manipulation can be directly supplied to make
interactive plots and Figure \ref{fig:interactive-linking} is an example
made with \pkg{plotly} and \pkg{crosstalk}. For example, the second row
results from selecting the highest mean maximum temperature in July and
this corresponds to Kalumburu where there is a larger winter (June to
August) temperature difference than summer (December to February). The
third row selects one of the stations (Birdsville Airport) with a large
average temperature difference in the interior desert region on the map
and the linked view from the time series shows that this maximum and
minimum temperature difference is constant throughout the year. The last
row selects the lowest maximum temperature in July on the temperature
band and surprisingly, the corresponding station doesn't fall into the
Tasmania island but Mountain Hotham in Victoria.

@@

Cubble is also compatible with \pkg{leaflet} and with leaflet extension
package \pkg{leafpop}, the temperature band can be displayed upon
clicking on the station on the map. Figure \ref{fig:interactive-popup}
shows more examples of temperature pattern that can be found in
Australia. With the same latitude as Kalumburu but in the Pacific
coastline, Cooktown airport shows a similarly high but constant
temperature throughout the year. With a slightly higher latitude than
the mountain Hotham, flat land in Melbourne airport shows a less extreme
temperature band. Lastly, the Cape Grim Baps, in the north-west tip of
the Tasmania island shows a uniformly low and thin temperature band.

@@

{\centering \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.5\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}

>>>>

To visualise climate variables from all the stations in Australia,
interactive graphics can be helpful. Here a plotly and a leaflet example
is given to explore the yearly diurnal temperature range pattern in
Australia. This would be useful to see how the average temperature range
differs in various latitudes in Australia, whether there's a summer or
winter effect, any coastline, mountain or desert topological
considerations that would affect the temperature.

Again, \texttt{weatherdata::climate\_full} will be first processed to
produce summarised monthly maximum and minimum temperature across 2016
to 2020. Then a shared data with the same group is created from the
nested and long cubble to enable the cross-talking via the common column
\texttt{id}. The map and temperature band is created separately with the
two shared data objects and combined together with
\texttt{crosstalk::bscols}. A sketch of the code to produce Figure
\ref{fig:interactive-linking} is presented below:

\begin{CodeChunk}
\begin{CodeInput}
R> # data processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # created grouped instance for crosstalk
R> nested <- clean %>% SharedData$new(~id, group = "cubble")
R> long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
R> 
R> # create the map with SharedData nested, the time series with SharedData long
R> p1 <- nested %>% ...
R> p2 <- long %>% ...
R> 
R> # Combine p1 and p2
R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
\end{CodeInput}
\end{CodeChunk}

In Figure \ref{fig:interactive-linking}, the first row shows the default
plot and when a selection is made, relevant component from both plots
will be highlighted. For example, in the second row, the highest mean
maximum temperature in July is selected on the time series and this
corresponds to Kalumburu on the map where there is a larger diurnal
temperature range in winter (June to August) than summer (December to
February). The third row selects a station (Birdsville Airport) in the
interior desert region that has a large average temperature difference
on the map and the linked view shows a wide but constant diurnal
temperature range throughout the year. The last row selects the lowest
maximum temperature in July on the right plot and surprisingly, this
links to the Thredbo airport at the boarder of Victoria and New South
Wales, rather than any station in the Tasmania island!

@@

With leaflet popup, the temperature range can be displayed as a small
plot upon clicking on the station on the map. This would require first
creating all the popup plot separately in a vector and then add these
plots to a leaflet map with \texttt{leafpop::addPopupGraphs()}:

\begin{CodeChunk}
\begin{CodeInput}
R> # create subplots for each station
R> df_id <- unique(clean$id)
R> p <- map(1:length(df_id), function(i){
+   dt <- clean %>% filter(id == df_id[i])
+   dt %>% ...
+ })
R> 
R> # inset subplots as a popup in leaflet
R> leaflet(clean) %>% 
+   addTiles() %>% 
+   addCircleMarkers(group = "a", ...) %>% 
+   leafpop::addPopupGraphs(graph = p, ...)
\end{CodeInput}
\end{CodeChunk}

Figure \ref{fig:interactive-popup} shows more examples of temperature
pattern that can be found in Australia. With the same latitude as
Kalumburu but in the Pacific coastline, Cooktown airport shows a
similarly high but constant temperature range throughout the year. With
a slightly higher latitude than the Thredbo airport , Melbourne airport
shows a less extreme temperature pattern Lastly, the Cape Grim Baps, in
the north-west tip of the Tasmania island shows low and narrow
temperature band, constantly betwee 10 and 20 degree throughout the
year.

@@

{\centering \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-north} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-vic} \includegraphics[width=0.45\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/popup-tas}
